---
title: "Chapter 3 Exercise 3.1"
author: "Alain Kuiete"
date: "6/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

3.1. The UC Irvine Machine Learning Repository6 contains a data set related
to glass identification. The data consist of 214 glass samples labeled as one
of seven class categories. There are nine predictors, including the refractive
index and percentages of eight elements: Na, Mg, Al, Si, K, Ca, Ba, and Fe.
The data can be accessed via:
```{r}
library(mlbench)
data(Glass)
str(Glass)
```

```{r}
head(Glass)
```


(a) Using visualizations, explore the predictor variables to understand their
distributions as well as the relationships between predictors.
(b) Do there appear to be any outliers in the data? Are any predictors skewed?
(c) Are there any relevant transformations of one or more predictors that
might improve the classification model?
 
We remove the target variable
```{r}
glass.predictors <- Glass[,-10]
```

Since all variables are numerical, we can use the skewness function of e1071 to estimated the predictors to represent.
```{r}
library(e1071)
skewValues <- apply(glass.predictors, 2, skewness)
skewValues
```

```{r}
par(mfrow = c(3,3))
hist(x = glass.predictors$RI)
hist(x = glass.predictors$Na)
hist(x = glass.predictors$Mg)
hist(x = glass.predictors$Al)
hist(x = glass.predictors$Si)
hist(x = glass.predictors$K)
hist(x = glass.predictors$Ca)
hist(x = glass.predictors$Ba)
hist(x = glass.predictors$Fe)
```

The predictors RI, Na, Al, Si and Ca are normal distributed
The predictors K, Ba, and Fe are rigtht skewed. we can aplied the log function on those variable to normalise or Boxcox to centralise, scale and transform.

The Mg predictor need to be centralise and scale. It is neither normal, nor skewed.

Looking for outliers
```{r}
par(mfrow = c(3,3))
boxplot(x = glass.predictors$RI, main = "RI")
boxplot(x = glass.predictors$Na, main = "Na")
boxplot(x = glass.predictors$Mg, main = "Mg")
boxplot(x = glass.predictors$Al, main = "Al")
boxplot(x = glass.predictors$Si, main = "Si")
boxplot(x = glass.predictors$K, main = "K")
boxplot(x = glass.predictors$Ca, main = "Ca")
boxplot(x = glass.predictors$Ba, main = "Ba")
boxplot(x = glass.predictors$Fe, main = "Fe")
```

```{r}
summary(glass.predictors)
```



To visualize the correlation between predictors, we use the corrplot function in the package of the same name.

```{r}
correlations <- cor(glass.predictors)
correlations
```

```{r}
library(corrplot)
corrplot(correlations, order = "hclust")
```


```{r}
GGally::ggpairs(as.data.frame(glass.predictors))
```

The only notable correlation is between RI and Ca.




















