---
title: "DAta 624 Group3 Project 2"
author: "Jeff Littlejohn"
date: "7/5/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

## Introduction

As most of us know, ensuring our beverages are produced at the correct potential for hydrogen (pH) level is an essential driver to our business. This pH score, the measure of acidity and alkalinity in our liquids, must be within a narrow, critical range to ensure long-term sales.

The objective of this project is to consider a number of measures and data points involved in the production of our beverages and build a model to use those factors to be able to predict the pH level of the beverage. 

Note that a factor being highly predictive of an outcome does not necessarily mean that the factor caused the outcome. For example, beverages that end up with higher than desired pH levels may also show high bowl setpoints. This does not mean that high bowl setpoints cause excessive pH values. They might both be similarly affected by an unknown cause, or large pH values may cause the high bowl setpoints. Our goal is simply to build a model that predicts pH levels. This process may lead to insights about possible causes of pH level problems, but that outcome is certainly not a given.

To construct this model, we had to first understand our data. We have more than 2800 production records of beverages that feature 33 data points, including Brand Code, Fill Ounces, PSC CO2, Temperature, etc. A full list of data elements will be conveyed in the next section. Some of the production records do not have values for all data elements, which is an obstacle we will had to overcome in building our model using established statistical principles. We also had to review

Data scientists use different methods to build models. No one approach gives the optimal approach for all data sets - and their underlying processes. To predict pH, we started by using linear regression, which constructs equations that look similar to you might have experienced in algebra at school. Next, we saw how effective nonlinear regression approaches such as neural networks and support vector machines (SVMs) are in predicting pH. Finally, we attempted to build models that use regression trees and rule-based models.

Don't be overwhelmed by mathematical terminology. We - or, rather our computers - are just using different methods to take the 33 different data elements and trying to use them to figure out their relationship to pH levels in our beverages.

```{r cars}
summary(cars)
```

## Approach

1. Exploratory Data Analysis
2. Data Processing
3. Linear Regression
4. Nonlinear Regression
5. Regression Trees and Rule-Based Models
6. Conclusion

## Exploratory Data Analysis

```{r}
bev_train <- readxl::read_excel('https://github.com/CerebralMastication/Presentations/raw/master/test.xlsx')
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
